### 2022-10-07 01_Species_abundance.Rmd
add NWSS data

---
title: "01_Vegetation_cover"
author: "Annegreet Veeken"
date: "24-11-2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(readxl)
library(LCVP)
library(lcvplants)
```

# To do

# Scotland

## QGIS

Import Scotland land cover HLCM2020_SCOTLAND_TIFF

-   Set layer style with .qml Import field site coordinate

-   Reproject to british national grid

Split vector layer based on Zone

-   Zone_A\_coordinates.shp = zone A coordinates

-   Zone_B\_coordinate.shp = zone B veg plot coordinates

Create buffer zone of 100 m and 1000 m

-   Zone_B\_buffer.shp

-   Zone_C\_buffer.shp

Use Zone histogram to compute pixels in zone attributed to certain land
cover type Export as Veg_cover_zoneC.csv and Veg_cover_zoneB.csv

Use Sample raster values to compute landcover types for vegetation
plots. Export as Veg_type_zoneB_sites.csv and Veg_type_zoneA_sites.csv,

Canopy data for zoneB was extracted for the zone B sites from the
Scottish Native Woodland survey:
<https://forestry.gov.scot/forests-environment/biodiversity/native-woodlands/native-woodland-survey-of-scotland-nwss>

Patterson, G., Nelson, D., Robertson, P. and Tullis, J., 2014.
*Scotland's Native Woodlands: Results from the Native Woodland Survey of
Scotland*. [online] Edinburgh: Forestry Commission Scotland. Available
at:
<https://forestry.gov.scot/publications/74-scotland-s-native-woodlands-results-from-the-native-woodland-survey-of-scotland>
[Accessed 3 October 2022].

-   Join NWSS woodland cover shape file with Species Structures table -
    by SCPTDATA_I

-   Union zone buffer (B or C separately) and NWSS woodland vector layer

-   remove all that don't overlap with the zone buffer

-   calculate area of each polygon with field calculator \$area

-   export to csv

## Zone A

```{r}
# load data
spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)
canop_a_raw <- read_xlsx("Data/Canopy_zone_A.xlsx")

canop_a <- canop_a_raw %>% 
  pivot_longer(-sitename, names_to = "species", values_to = "abun") %>% 
  mutate(direction = "Canopy", distance = "0-10 meter")
# Make tidy
spec_a <- spec_a_raw %>% 
  pivot_longer(3:19, names_to = "plot", values_to = "abun") %>% 
  rename(sitename = Site, species = Species) %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  # create direction and distance column
  mutate(direction = case_when(str_detect(plot, "N") ~ "North",
                               str_detect(plot, "S") ~ "South",
                               str_detect(plot, "E") ~ "East",
                               str_detect(plot, "W") ~ "West",
                               str_detect(plot, "NE") ~ "Northeast",
                               str_detect(plot, "NW") ~ "Northwest",
                               str_detect(plot, "SW") ~ "Southwest",
                               str_detect(plot, "SE") ~ "Southeast",
                               plot == "Center" ~ "Center"),
         distance = case_when(plot %in% c("N...4", "E...5", "S...6","W...7") ~ 
                                "1.5-3 meter",
                              plot %in% c("N...8", "E...9", "S...10", "W...11")  ~ 
                                "3-6 meter",
                              plot %in% c("N...12","NE","SE","NW","SW", "E...14", "S...16", "W...18") ~ 
                                "6-10 meter",
                              plot == "Center" ~ "0 meter")) %>% 
  bind_rows(canop_a) %>% 
  # create unique id for plot in site
  group_by(sitename, direction, distance) %>% 
  mutate(plot = cur_group_id())

```

## Zone B & C

Vegetation cover data.

```{r load and prepare veg cover data}
# load data
zone_b <- read.csv("Data/Veg_cover_zoneB.csv")
zone_c <- read.csv("Data/Veg_cover_zoneC.csv")

# EUNIS land cover classifications
veg_type_labels <- c(HISTO_1 = "C1: Water",
                     HISTO_2 = "D1: Raised and blanket bogs",
                     HISTO_3 = "D2: Valley mires, poor fens and transition mires",
                     HISTO_4 = "D4: Base-rich fens and calcareous spring mires", 
                     HISTO_5 = "E1: Dry grasslands",
                     HISTO_6 = "E2: Mesic grasslands",
                     HISTO_7 = "E3: Seasonally wet and wet grasslands",
                     HISTO_8 = "E4: Alpine and subalpine grasslands",
                     HISTO_9 = "E5: Woodland fringes and clearings and tall forb stands",
                     HISTO_10 = "F2: Arctic, alpine and subalpine scrub",
                     HISTO_11 = "F3: Temperate and mediterranean-montane scrub",
                     HISTO_12 = "F4: Temperate shrub heathland",
                     HISTO_13 = "F9: Riverine and fen scrubs",
                     HISTO_14 = "G1: Broadleaved deciduous woodland",
                     HISTO_15 = "G3: Coniferous woodland",
                     HISTO_16 = "G4: Mixed deciduous and coniferous woodland",
                     HISTO_17 = "G5: Line of trees, amall anthropogenic woodlands, recently felled woodland, early-stage woodland and coppice",
                     HISTO_18 = "H2: Screes",
                     HISTO_19 = "Inland cliffs, rock pavements and outcrops",
                     HISTO_20 = "I1: Arable land and market gardens",
                     HISTO_21 = "J: Builtup (and Bare before post)",
                     HISTO_23 = "O: Bare (after post)")

# make data tidy
zone_b <- zone_b %>% 
  dplyr::select(sitename = name, lat, lon, contains("HISTO")) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = recode(veg_type, !!!veg_type_labels),
  # rename sitenames       
         sitename = str_remove(sitename, "-A")) %>% 
  # filter land cover type that are not vegation
  filter(!veg_type %in% c("C1: Water", 
                         "J: Builtup (and Bare before post)",
                         "O: Bare (after post)")) %>% 
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov_b = value/sum(value)) %>%
  group_by(sitename, veg_type) %>%
  summarise(veg_cov_b = sum(veg_cov_b, na.rm = TRUE))

# make data tidy
zone_c <- zone_c %>% 
  dplyr::select(sitename = name, lat, lon, contains("HISTO")) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = recode(veg_type, !!!veg_type_labels),
  # rename sitenames       
         sitename = str_remove(sitename, "-A")) %>% 
  # filter land cover type that are not vegation
  filter(!veg_type %in% c("C1: Water", 
                         "J: Builtup (and Bare before post)",
                         "O: Bare (after post)")) %>% 
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov_c = value/sum(value)) %>%
  group_by(sitename, veg_type) %>%
  summarise(veg_cov_c = sum(veg_cov_c, na.rm = TRUE))

veg_type <- left_join(zone_b, zone_c, by = c("sitename", "veg_type")) 
```

```{r load and prepare species data}
# Load in vegetation type data for the plots
zone_b_site <- read.csv("Data/20220106-Veg_type_zoneB_sites.csv")
spec_b_raw <- read_xlsx("Data/species_zone_B.xlsx") # load in species data

# Make vegetation type tidy
zone_b_site <- zone_b_site %>% 
  # create seperate site and plot column
  separate(name, c("sitename", "plot")) %>% 
  # recode veg_type
  mutate(veg_type = as.factor(rvalue_1)) %>% 
  dplyr::select(sitename, plot, lat, lon, veg_type) %>% 
  mutate(veg_type = recode(veg_type, 
                           "1" = "C1: Water",
                           "2" = "D1: Raised and blanket bogs",
                           "3" = "D2: Valley mires, poor fens and transition mires",
                           "4" = "D4: Base-rich fens and calcareous spring mires", 
                           "5" = "E1: Dry grasslands",
                           "6" = "E2: Mesic grasslands",
                           "7" = "E3: Seasonally wet and wet grasslands",
                           "8" = "E4: Alpine and subalpine grasslands",
                           "9" = "E5: Woodland fringes and clearings and tall forb stands",
                           "10" = "F2: Arctic, alpine and subalpine scrub",
                           "11"  = "F3: Temperate and mediterranean-montane scrub",
                           "12" = "F4: Temperate shrub heathland",
                           "13" = "F9: Riverine and fen scrubs",
                           "14" = "G1: Broadleaved deciduous woodland",
                           "15" = "G3: Coniferous woodland",
                           "16" = "G4: Mixed deciduous and coniferous woodland",
                           "17" = "G5: Line of trees, amall anthropogenic woodlands, recently felled woodland, early-stage woodland and coppice",
                           "18" = "H2: Screes",
                           "19" = "Inland cliffs, rock pavements and outcrops",
                           "20" = "I1: Arable land and market gardens",
                           "21" = "J: Builtup (and Bare before post)",
                           "23" = "O: Bare (after post)"))

# Tidy and join 
spec_b <- spec_b_raw %>% 
  pivot_longer(contains("plot"), names_to = "plot", values_to = "abun",
               values_drop_na = TRUE) %>% 
  mutate(plot = str_remove(plot, "plot_ ")) %>% 
  dplyr::select(sitename = Site, plot, species = Species, abun) %>% 
  # join with vegetation type data
  left_join(zone_b_site, by = c("sitename", "plot")) %>% 
  # filter trees from species survey (will be covered by NWSS)
  filter(!species %in% c("Betula pendula", "Pinus sylvestris", 
                        "Juniperus communis", "Salix sp.",
                        "Larix decidua", "Salix repens"))
```

tree abundance

```{r}
canop_b_raw <- 
  read.csv("Data/NWSS_zoneB_ID_forest_cover.csv") %>% 
  # clean column names
  rename_with(~str_remove(., "NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"),
              matches("NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"))
canop_c_raw <- 
  read.csv("Data/NWSS_zoneC_ID_forest_cover.csv") %>% 
  # clean column names
  rename_with(~str_remove(., "NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"),
              matches("NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"))
NWSS_sp <- 
  read.csv("Data/NWSS_Species_Structures.csv") 
NWSS <- 
  read.csv("Data/NWSS_basis.csv") %>%  # inventory of NWSS 
  # clean column names
  rename_with(~str_remove(., "NWSS_Species_Structures_"),
              matches("NWSS_Species_Structures_"))

# calculate average species abundance of species per habitat type
# species percentage of NWSS (not available for all polygon)
habitats <- canop_c_raw$DOM_HABITA %>% unique
sp_habitat <- NWSS %>% 
  filter(DOM_HABITA %in% habitats) %>% 
  group_by(DOM_HABITA, SPECIES) %>% 
  summarise(mean_pct = mean(SP_STR_PCT))

# - calculate woodland cover for every site
# - weight woodland cover by mean abundance of species per type
canop_b <- canop_b_raw %>% 
  # select relevant columns
  dplyr::select(sitename = name, SCPTDATA_I, TYPE, DOM_HABITA, woodland_area = Woodland_area) %>% 
  # fill in area without woodland
  mutate(TYPE = if_else(TYPE == "", "Open land habitat", TYPE),
         # rename site names
         sitename = str_remove(sitename, "-A")) %>% 
  # calculate woodland type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(total_area = sum(woodland_area)) %>% 
  mutate(woodland_percentage = woodland_area/total_area) %>% 
  # join with species abundance
  left_join(sp_habitat, by = "DOM_HABITA") %>% 
  # create column with scientific species name
  mutate(species = recode(SPECIES, "Common alder*" = "Alnus glutinosa",
                          "Rowan *" = "Sorbus aucuparia",
                          "Sitka spruce" = "Picea sitchensis",
                          "European larch" = "Larix decidua",
                          "Lodgepole pine" = "Pinus contorta",
                          "Downy birch *" = "Betula pubescens",
                          "Scots pine *" = "Pinus sylvestris",
                          "Hazel *" = "Corylus avellana",
                          "Eared willow *+" = "Salix aurita", 
                          "Grey willow *" = "Salix cinerea", 
                          "Hybrid larch" = "Larix decidua", 
                          "Shrubs - preferentially browsed +" = "Unidentified",
                          "Dwarf juniper *+" = "Juniperus communis", 
                          "Silver birch *" = "Betula pendula", 
                          "Beech" = "Fagus sylvatica", 
                          "Lawsons cypress" = "Chamaecyparis lawsoniana", 
                          "Common alder*" = "Alnus glutinosa", 
                          "Douglas fir" = "Pseudotsuga menziesii", 
                          "Yew *" = "Taxus baccata", 
                          "Upright juniper * +" = "Juniperus communis", 
                          "Norway spruce" = "Picea abies", 
                          "Bird cherry *" = "Prunus padus", 
                          "Intermediate oak *" = "Quercus sp.", 
                          "Sessile oak *" = "Quercus petraea", 
                          "Ash *" = "Fraxinus excelsior", 
                          "Holly *" = "Ilex aquifolium", 
                          "Aspen *" = "Populus tremula", 
                          "Goat willow*" = "Salix caprea", 
                          "Elder *" = "Sambucus nigra", 
                          "Broom *+" = "Cytisus scoparius", 
                          "Other non-native trees/shrubs" = "Unidentified", 
                          "Sycamore" = "Acer pseudoplatanus", 
                          "Grand fir" = "Abies grandis", 
                          "Gorse *+" = "Ulex europaeus", 
                          "Gean wild cherry *" = "Prunus avium",
                          "Austrian pine" = "Pinus nigra",
                          "Blackthorn *+" = "Prunus spinosa",
                          "Birch *" = "Betula sp.",
                          "Coast redwood" = "Sequoia sempervirens",
                          "Bramble *+" = "Rubus fruticosus",
                          "Crab apple" = "Malus sylvestris",
                          "Dog rose *+" = "Rosa canina",
                          "Geulder rose *+" = "Viburnum opulus",
                          "Grey alder" = "Alnus incana",
                          "Hawthorn *" = "Crataegus monogyna",
                          "Field maple" = "Acer campestris",
                          "Hornbeam" = "Carpinus betulus",               
                          "Horse chestnut" = "Aesculus hippocastanum",
                          "Japanese knotweed +" = "Reynoutria japonica",
                          "Japanese larch" = "Larix kaempferi",
                          "Leyland cypress" = "Cupressus sp.",        
                          "Lime all" = "Tilia sp.",
                          "Noble fir" = "Abies procera",
                          "Non-native willow" = "Salix sp.",
                          "Norway maple" = "Acer platanoides",       
                          "Nothofagus all" = "Nothofagus sp.",
                          "Omorika spruce" = "Picea omorika",         
                          "Other native willows *" = "Salix sp.",
                          "Other pines" = "Pinus sp.",
                          "Other whitebeam"  = "Sorbus aria",       
                          "Pedunculate oak *" = "Quercus robur",     
                          "Poplar" = "Populus sp.",
                          "Red oak" = "Quercus rubra",               
                          "Rhododendron ponticum +" = "Rhododendron ponticum",
                          "Scots pine" = "Pinus sylvestris",
                          "Silver fir" = "Abies alba",
                          "Snowberry+" = "Symphoricarpos sp.",
                          "Sweet chestnut" = "Castanea sativa",
                          "Western hemlock" = "Tsuga heterophylla",
                          "Western red cedar" = "Thuja plicata",
                          "Wild service tree" = "Sorbus torminalis",      
                          "Wych elm *" = "Ulmus glabra",
                          "English elm" = "Ulmus procera",
                          "Japanese cedar" = "Cryptomeria japonica",
                          "Montane willow all *+" = "Salix sp."
                          )) %>% 
  # filter non trees
  mutate(growthform = case_when(str_detect(SPECIES, "\\+") ~ "shrub",
                                SPECIES %in% c("Japanese knotweed +") ~ "herb",
                                TRUE ~ "tree")) %>% 
  filter(growthform == "tree",
         !species %in% c("Unidentified","")) %>% 
  # weight species abundance by woodland percentage
  group_by(sitename, species) %>% 
  mutate(sp_abun = mean_pct * woodland_percentage) %>% 
  summarise(spec_abun_b = sum(sp_abun))

# PAWS =  Planted Woods on Ancient Woodland Sites
canop_c <- canop_c_raw %>% 
  # select relevant columns
  dplyr::select(sitename = name, SCPTDATA_I, TYPE, DOM_HABITA, woodland_area = Woodland_area) %>% 
  # fill in area without woodland
  mutate(TYPE = if_else(TYPE == "", "Open land habitat", TYPE),
         # rename site names
         sitename = str_remove(sitename, "-A")) %>% 
  # calculate woodland type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(total_area = sum(woodland_area)) %>% 
  mutate(woodland_percentage = woodland_area/total_area) %>% 
  # join with species abundance
  left_join(sp_habitat, by = "DOM_HABITA") %>% 
  # create column with scientific species name
  mutate(species = recode(SPECIES, "Common alder*" = "Alnus glutinosa",
                          "Rowan *" = "Sorbus aucuparia",
                          "Sitka spruce" = "Picea sitchensis",
                          "European larch" = "Larix decidua",
                          "Lodgepole pine" = "Pinus contorta",
                          "Downy birch *" = "Betula pubescens",
                          "Scots pine *" = "Pinus sylvestris",
                          "Hazel *" = "Corylus avellana",
                          "Eared willow *+" = "Salix aurita", 
                          "Grey willow *" = "Salix cinerea", 
                          "Hybrid larch" = "Larix decidua", 
                          "Shrubs - preferentially browsed +" = "Unidentified",
                          "Dwarf juniper *+" = "Juniperus communis", 
                          "Silver birch *" = "Betula pendula", 
                          "Beech" = "Fagus sylvatica", 
                          "Lawsons cypress" = "Chamaecyparis lawsoniana", 
                          "Common alder*" = "Alnus glutinosa", 
                          "Douglas fir" = "Pseudotsuga menziesii", 
                          "Yew *" = "Taxus baccata", 
                          "Upright juniper * +" = "Juniperus communis", 
                          "Norway spruce" = "Picea abies", 
                          "Bird cherry *" = "Prunus padus", 
                          "Intermediate oak *" = "Quercus sp.", 
                          "Sessile oak *" = "Quercus petraea", 
                          "Ash *" = "Fraxinus excelsior", 
                          "Holly *" = "Ilex aquifolium", 
                          "Aspen *" = "Populus tremula", 
                          "Goat willow*" = "Salix caprea", 
                          "Elder *" = "Sambucus nigra", 
                          "Broom *+" = "Cytisus scoparius", 
                          "Other non-native trees/shrubs" = "Unidentified", 
                          "Sycamore" = "Acer pseudoplatanus", 
                          "Grand fir" = "Abies grandis", 
                          "Gorse *+" = "Ulex europaeus", 
                          "Gean wild cherry *" = "Prunus avium",
                          "Austrian pine" = "Pinus nigra",
                          "Blackthorn *+" = "Prunus spinosa",
                          "Birch *" = "Betula sp.",
                          "Coast redwood" = "Sequoia sempervirens",
                          "Bramble *+" = "Rubus fruticosus",
                          "Crab apple" = "Malus sylvestris",
                          "Dog rose *+" = "Rosa canina",
                          "Geulder rose *+" = "Viburnum opulus",
                          "Grey alder" = "Alnus incana",
                          "Hawthorn *" = "Crataegus monogyna",
                          "Field maple" = "Acer campestris",
                          "Hornbeam" = "Carpinus betulsu",               
                          "Horse chestnut" = "Aesculus hippocastanum",
                          "Japanese knotweed +" = "Reynoutria japonica",
                          "Japanese larch" = "Larix kaempferi",
                          "Leyland cypress" = "Cupressus sp.",        
                          "Lime all" = "Tilia sp.",
                          "Noble fir" = "Abies procera",
                          "Non-native willow" = "Salix sp.",
                          "Norway maple" = "Acer platanoides",       
                          "Nothofagus all" = "Nothofagus sp.",
                          "Omorika spruce" = "Picea omorika",         
                          "Other native willows *" = "Salix sp.",
                          "Other pines" = "Pinus sp.",
                          "Other whitebeam"  = "Sorbus aria",       
                          "Pedunculate oak *" = "Quercus robur",     
                          "Poplar" = "Populus sp.",
                          "Red oak" = "Quercus rubra",               
                          "Rhododendron ponticum +" = "Rhododendron ponticum",
                          "Scots pine" = "Pinus sylvestris",
                          "Silver fir" = "Abies alba",
                          "Snowberry+" = "Symphoricarpos sp.",
                          "Sweet chestnut" = "Castanea sativa",
                          "Western hemlock" = "Tsuga heterophylla",
                          "Western red cedar" = "Thuja plicata",
                          "Wild service tree" = "Sorbus torminalis",      
                          "Wych elm *" = "Ulmus glabra",
                          "English elm" = "Ulmus procera",
                          "Japanese cedar" = "Cryptomeria japonica",
                          "Montane willow all *+" = "Salix sp.",
                          "Arran whitebeam *" = "Sorbus arranensis"
         )) %>% 
  # filter non trees
  mutate(growthform = case_when(str_detect(SPECIES, "\\+") ~ "shrub",
                                SPECIES %in% c("Japanese knotweed +") ~ "herb",
                                TRUE ~ "tree")) %>% 
  filter(growthform == "tree",
       !species %in% c("Unidentified","")) %>% 
  # weight species abundance by woodland percentage
  group_by(sitename, species) %>% 
  mutate(sp_abun = mean_pct * woodland_percentage) %>% 
  summarise(spec_abun_c = sum(sp_abun))

# join df's from both zones
canop <- canop_b %>% 
  left_join(canop_c, by = c("sitename","species"))
```

```{r calculate vegetation cover}
veg_cover_bc <- spec_b %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  # calculate average species cover per vegetation type using the plots belonging to that vegetation type
  group_by(sitename, veg_type, species) %>% 
  summarise(abun = mean(abun)) %>% 
  # join with vegetation cover data
  full_join(veg_type, by = c("sitename", "veg_type"))

# print table with missing species data for vegetation types
missing_vegtype <- veg_cover_bc %>% 
  filter(is.na(species)) %>% 
  group_by(sitename, veg_type) %>% 
  summarise(cov_b = round(sum(veg_cov_b), 2), cov_c = round(sum(veg_cov_c), 2)) %>%
  filter(!cov_b == 0 | !cov_c == 0)

# for missing data, use species data from other plots
missing <- spec_b %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  filter(veg_type %in% missing_vegtype$veg_type) %>% 
  group_by(veg_type, species) %>% 
  summarise(abun = mean(abun))
veg_cover_bc_na <- veg_cover_bc %>% 
  filter(is.na(species)) %>%
  dplyr::select(-species, -abun) %>% 
  left_join(missing, by = "veg_type")

veg_cover_bc <-  veg_cover_bc %>% 
  filter(!is.na(species)) %>% 
  # bind with df with missing veg types
  bind_rows(., veg_cover_bc_na) %>% 
  # weight species abundance by veg cover
  group_by(sitename, species) %>% 
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>%
  summarise(spec_abun_b = sum(spec_abun_b),
            spec_abun_c = sum(spec_abun_b)) %>% 
  # bind with tree species data
  bind_rows(canop) %>% 
  arrange(sitename, species)
```

## Standardization species names

```{r standardize species names}
# load in generalized species list, spelling checked and harmonized with LCVP
harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") 
harm <- harm %>% 
  mutate(stand.spec = str_remove(stand.spec, pattern = " sp\\."))

spec_a <- spec_a %>% 
  left_join(harm, by = "species")
veg_cover_bc <- veg_cover_bc %>% 
  left_join(harm, by = "species") %>% 
  filter(!is.na(stand.spec))

saveRDS(veg_cover_bc, "RDS_files/01_Species_abundance_bc.rds")
saveRDS(spec_a, "RDS_files/01_Species_abundance_a.rds")
```

# Switzerland

## QGIS

Land cover data was downloaded from:
<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability.html>

To decrease file size, I only selected relevant columns: the one
containing coordinates and land cover data using the following code:

```{r eval = FALSE}
# ag_b_00_03_37_area_csv <- read_delim(
#     "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data analysis/PhD/Validation-study/Data/Swiss_LC/ag-b-00.03-37-area-csv.csv",+delim = ";",
#     escape_double = FALSE,
#     trim_ws = TRUE
#   )
# 
# LC <- ag_b_00_03_37_area_csv %>%
#   dplyr::select(RELI, E_COORD = E, N_COORD = N, LC18_6, LC18_27)
# 
# write.csv(LC, file = "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data
#           analysis/PhD/Validation-study/Data/Swiss_LC/Swiss_LC18.csv")

```

In this file I found the CRS:
<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability/swiss-land-use-statistics.assetdetail.20104775.html>\
"Les coordonnées « E_COORD » et « N_COORD » des hectares dans le système
de coordonnées suisse MN95 (7 chiffres). Les coordonnées représentent le
point d'échantillonnage correspondant (intersections des coordonnées
hectométriques)."

Import Swiss land cover (100m resolution) Swiss_LC18.csv

-   Set CRS code to EPSG:4150 Rasterize this layer using the GDAL
    rasterize (vector to raster) tool
-   Set burnin to LC18_27
-   Horizontal and vertical resolution to 100 m
-   export and set resolution to 25 m for merge with forest cover data

Import forest cover data in ArcGIS pro reproject to EPSG:4150 and export
raster(folder called reproj)/ Import to QGIS set layer crs to EPSG:4150

Import coordinates of pollen plots (Pollen_site_coord.csv).

Create buffer zone of 100 m and 1000 m

Use Zonal histogram to compute pixels in zone attributed to certain land
cover type Export as Veg_cover_zoneC.csv and Veg_cover_zoneB.csv

Import coordinate of all BDM sites Use Sample raster values to compute
landcover types for vegetation plots. Export as Veg_type_plots_csv

```{r}
zone_b_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneB_swiss.csv")
zone_c_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneC_swiss.csv")
LC_labels <- read_xlsx("Data/Swiss_LC/be-b-00.03-37-area-var-v01.xlsx",
                       sheet = "Codes LC_27")
FC_labels <- read.csv("Data/Swiss_LC/Forest_cover_labels.csv") %>% 
  rename(CODE = category)
LC_labels <- LC_labels %>% 
  dplyr::select(CODE, LABEL_DE) %>% 
  bind_rows(FC_labels)

zone_b <- zone_b_raw %>% 
  # make tidy
  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
  filter(!value == 0) %>% 
  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
  # filter land cover type that are not vegation
  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
                           "Versteinte Flächen","Befestigte Flächen",
                           "Gemischte Kleinstrukturen")) %>%
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov = value/sum(value)) %>%
  group_by(sitename, LABEL_DE) %>%
  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))

zone_c <- zone_c_raw %>% 
  # make tidy
  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
  filter(!value == 0) %>% 
  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
  # filter land cover type that are not vegation
  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
                           "Versteinte Flächen","Befestigte Flächen",
                           "Gemischte Kleinstrukturen")) %>%
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov = value/sum(value)) %>%
  group_by(sitename, LABEL_DE) %>%
  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))

# merge zone b and c
veg_cover <- zone_b %>% 
  full_join(zone_c, by = c("sitename", "LABEL_DE"), 
            suffix = c("_b","_c")) %>% 
  replace_na(list(veg_cov_b = 0))
```

```{r join veg type data and species data}
# Veg type of BDM sites
veg_type <- read.csv("Data/Swiss_LC/FC_LC_type_swiss.csv")

# Join with labels
veg_type <- veg_type %>% 
  left_join(LC_labels, by = c("SAMPLE_1" = "CODE"))

# Load in species data
bdm_sp <- readRDS("RDS_files/01_Species_abundance_Swiss.rds")

bdm_sp <- bdm_sp %>% 
  # filter stuff that is not a plant
  filter(!stand.spec == "No plants") %>% 
  # filter unindentified
  filter(!stand.spec == "Unindentified") %>% 
  # join with vegetation type data
  left_join(veg_type, by = c("sitename" = "aID_STAO")) 
# create table with genus and family to join after summarizing
genus_fam <- bdm_sp %>% 
  ungroup() %>% 
  dplyr::select(family, genus, stand.spec) %>% 
  distinct
```

```{r}
spec_cover <- bdm_sp %>% 
  # calculate average species cover per vegetation type using the plots belonging to that vegetation type
  group_by(sitename, LABEL_DE, stand.spec) %>% 
  summarise(abun = mean(abun)) %>% 
  # join with vegetation cover data
  right_join(veg_cover, by = c("sitename", "LABEL_DE")) %>% 
  # weight species abundance by veg cover
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c)

# print table with missing species data for vegetation types
missing_vegtype <- spec_cover %>% 
  filter(is.na(stand.spec)) %>% 
  group_by(sitename, LABEL_DE) %>% 
  summarise(cov_b = round(sum(veg_cov_b), 2), 
            cov_c = round(sum(veg_cov_c), 2)) %>%
  filter(!cov_b == 0 |!cov_c == 0)

# for missing data, use species data from other plots
missing <- bdm_sp %>% 
  filter(LABEL_DE %in% missing_vegtype$LABEL_DE) %>% 
  group_by(LABEL_DE, stand.spec) %>% 
  summarise(abun = mean(abun))
spec_cover_na <- spec_cover %>% 
  filter(is.na(stand.spec)) %>%
  dplyr::select(sitename, LABEL_DE, veg_cov_b, veg_cov_c) %>% 
  left_join(missing, by = "LABEL_DE") %>% 
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) 
spec_cover_bc <-  spec_cover %>% 
  filter(!is.na(stand.spec)) %>%
  bind_rows(., spec_cover_na) %>% 
  arrange(sitename, LABEL_DE, stand.spec) %>% 
  dplyr::select(sitename, stand.spec, spec_abun_b, spec_abun_c) %>% 
  ungroup() %>% 
  left_join(genus_fam, by = "stand.spec")

saveRDS(spec_cover_bc, "RDS_files/01_Species_abundance_zoneBC_swiss.rds")
```