### 2022-12-28 Manuscript.Rmd
method adjustments

---
output:
  word_document:
    reference_docx: "Manuscript-Markdown-template.docx"
    fig_caption: yes
    bibliography: "Validation_study_references.bib" 
---

```{r setup, include=FALSE}
# Document options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      collapse = TRUE,
                      results = "hide")

source("01_Prepare_pollen_data.R")

# Libraries
library(bibtex)
library(knitr)

# Datasets
bdm_meta <- readRDS("RDS_files/01_Meta_data_BDM.rds")
```

(working title)

# Reconstructing plant trait composition from palaeoecological data: testing the validity of the use of pollen data for studying long-term changes in plant trait composition

## Abstract

-   needs more testing

-   we advocate for more testing with greater sample size

-   need to take in to account trait variation pollen types

## Introduction

Functional traits provide a better insight in to the processes that drive biodiversity and how biodiversity affects ecosystem functioning [@Hevia_2017]. While the functioning of modern ecosystems is regularly contingent on past human and environmental factors [@Levis_2018; @Jogiste_2017; @Blonder_2018], the observational duration of trait-based studies is generally at most a couple of decades [@Estes_218]. So to extend the time scales of ecosystem functioning research, considerable interest in the trait-based approach has arisen in palaeoecology [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019]. Applying the trait-based approach in palaeoecological studies could offer a new way of interpreting pollen data and increase our understanding of the drivers of vegetation change on time-scales of centuries to millennia. For instance, @Sande_2019 found that on a 7000-year time scale human disturbance supersedes climate effects on plant trait composition in the Amazon.

Although palaeoecological studies that have applied the trait-based approach show promising results, it is a relatively new approach of which the intricacies are under-explored. Little is known about how representative the pollen record is of past plant trait composition. Two main methodological issues arise when attempting to reconstruct functional trait composition from pollen data. First, pollen quantity does not exactly represent past plant abundance. Second, the low taxonomic resolution of the pollen record makes it not straightforward to assign traits to pollen taxa, as trait data is generally collected at the species level.

Taxonomic differences in pollen dispersal and production can cause an over- or under-representation of taxa in the pollen record. Wind-pollinated and woody taxa are generally better represented in the pollen record than herbaceous and insect-pollinated taxa as their pollen are produced in larger quantities and dispersed over greater distances. The source area of a pollen of a wind-pollinated taxon can therefore be a couple magnitudes greater than that of a insect-pollinated taxon [@Meltsov_2011] (extend argument? bunting 2003, bunting 2022). Over- or underrepresentation can be corrected for with the use of taxon-specific pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022]. Furthermore, it has been suggested by @Blaus_2020 to treat woody and non-woody taxa separately in the reconstruction of functional composition from pollen data as their representative pollen source area differs greatly.

Because of the difference in  pollen productivity and dispersal among pollen taxa, pollen counts might not reflect absolute abundance and hence trait composition [@Sande2021]. Alternative to forming a direct representation of past plant trait composition, pollen based trait reconstruction might be better suited to study relative changes in trait composition. For instance, @Sande2021 showed that pollen assemblages reflect relative trait changes in trait composition of the vegetation along an elevation gradient.

The second challenge in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020]. Trait variation within taxa can be substantial depending on the trait and the diversity of the taxon [@Albert_2010; @Prinzing_2008]. It can be expected that this is especially problematic in the case of large, abundant plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family. @Blaus_2020 and @Carvalho_2019 proposed that the high species number included in herbaceous pollen taxon could be one of the reasons why herbaceous vegetation was less well represented by pollen based functional diversity reconstructions. In this study, we use a Bayesian approach for modelling trait composition from pollen data, thereby allowing for the propagation of uncertainty in the taxon level trait reconstruction to the community level.

With this study, we aim to test the validity of using pollen records for past plant trait composition reconstruction. We address two questions. Firstly, what factors influence the pollen - vegetation trait relationship? Using a data set from the Cairngorms, Scotland, we investigate the effect of spatial scale, plant functional group, correction factors and taxonomic resolution on the representativeness of the pollen assemblages for functional trait reconstruction. Secondly, can pollen based trait reconstruction predict changes in trait composition along environmental gradients? Using pollen, vegetation and trait data collected along an elevation gradient in Switzerland we test if pollen assemblages can predict relative changes in trait composition with elevation.

## Method

### Study sites
We collected pollen, trait and species composition data from 16 sites in the Cairngorms National Park, Scotland, United Kingdom. We selected sites to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest. Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other. Pollen, trait and species data was also available from `r nrow(bdm_meta)` sites across Switzerland were monitored in the context of the Biodiversity Monitoring Switzerland (BDM). This is a systematic sampling grid for the long-term monitoring of plant and animal species in Switzerland [@BDM_2022]. Trait and species composition data, but not pollen data, from an additional `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - nrow(bdm_meta)` sites that were surveyed in this network were also used in this study.

### Study design
We used the Scottish data to test if pollen-based trait reconstructions represents vegetation trait composition. We compare the community weighted mean (CWM) trait value of the vegetation with the CWM trait values reconstructed from surface pollen samples. Study design mostly followed the protocol of @Bunting_2013, which allows for the calculation of vegetation composition in concentric circles around the moss sampling point to test the spatial scale that is represented by the moss sample (Appendix). We calculated the vegetation composition in three nested concentric rings of 10 m, 100 m and 1000 m in radius. The inner ring (10 m) and the middle ring was surveyed in the field. The vegetation cover of the outer ring (1000 m) was estimated using QGIS and land cover maps [\@scotisch_government; \@QGIS_software]. To test the effect of spatial scale on the representativeness of pollen assemblages for trait composition reconstruction, we compare vegetation CWM and pollen CWM in the rings of 10 m, 100 m and 1000 m. Furthermore, four different treatments were assigned to test what factors could influence representation of trait composition by pollen data. Firstly, the use of correction factors, than the separation of vegetation and pollen data in to plant functional type and pollination mode. Lastly, the vegetation was aggregated to genus level and family level to test the effect of taxonomic resolution on the reconstructions.

To test if pollen-trait reconstructions can predict relative changes in vegetation trait composition along environmental gradients, we calculate changes in CWM trait values along an elevation gradient in Switzerland. The sites span an elevation between `r min(bdm_meta$elevation)` m and `r max(bdm_meta$elevation)`m. Of these sites `r table(bdm_meta$habitat01)[1]` were in forests and `r table(bdm_meta$habitat01)[2]` in grassland.

### Vegetation survey
At the Scottish sites, we surveyed 17 sub-plots of 0.25 m^2^ in the inner ring (10 m): one in the center, 3 in every wind direction at 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix). In the middle ring (10-100 m), we selected between 4 and 6 plots of 5 by 5 meter based on a stratified random design based on vegetation types that were identified in the fields. Vegetation abundance was estimated using percentage cover. In Switzerland, the plots were 1.8 m in radius and contained 4 sub-plots. Each sub-plot was a quadrant of 0.16 m^2^ with nine points (a 3 x 3 grid), and percentage coverage of the species present at the nine points was estimated. All species names were standardized according to the Leipzig Plant Catalogue using the *lcvplants* and *lcvp* R packages [@freiberg2020lcvp].

### Trait data collection
We used three traits for the analysis, specific leaf area (SLA), leaf area (LA) and and plant height. These traits reflect different components in the spectrum of plant strategies [@Diaz2016]. In Scotland, we collected trait data in the middle ring (100 m). For determining plant height, we randomly selected 10 flowering individuals of each species at every site. Non-flowering individuals were selected if flowering individuals were not available. For determining specific leaf area and leaf area, we collected five leaf samples per species in sealable plastic bags. Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner. Samples were dried at 70 ***°***C for 72 hours to constant dry weight before measuring dry weight. In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same. For reconstructing trait composition from pollen data, we did not use the data collected in the field but previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].

### Pollen data collection
Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Pardoe2010]. Moss was collected from multiple location within a 1 m radius in the center of each plot.Pollen were extracted from the moss using standardized pollen preparation protocol (REF). Pollen were identified using light microscope and the key of (REF).

For reconstructing plant trait composition from pollen data, assumptions need to be made about the species belonging to the pollen taxa in the data set. Therefore we created a pollen to species translation table. We used the *rgbif* package to download the current distribution of species belonging to each pollen taxon in Scotland and Switzerland from the Global Biodiversity Information Facility database (REF). Species distribution data was queried using the pollen taxonomic name, i.e. the lowest taxonomic name the pollen could be identified to. All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level.

### Calculation of community weighted mean
We use the community weighted mean (CWM) to characterize functional composition. Here CWM is specified as the trait distribution of a taxon, i.e. the species in the vegetation data and the pollen taxon in the pollen data, weighted by its abundance. First, the trait distribution on the species or the pollen taxonomic level was estimated. We assume that the trait data of SLA, LA and plant height follow a lognormal distribution for the likelihood, so that
$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$
Where *i* represents each trait observation of a taxon (*j*) from the trait data. $\mu_{j}$ and $\sigma_{j}$ are the mean trait value and the standard deviation per taxon, respectively. Vague priors were used for the taxon mean and standard deviation estimation. We then modeled the CWM as a mixture of normal distributions that were weighted by the species abundance or pollen percentage, so that

$$
CWM_{s} \sim \sum^{n}\pi_{j}Normal(\mu_{j},\sigma_{j})
$$

Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon or species at a given site *s*. The mean ($CWM_{mean}$) and the standard deviation ($CWM_{sd}$) of this modeled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.

### Relationship between pollen and vegetation community weighted mean
To test the relationship between the pollen based reconstruction of CWM and the vegetation CWM, we standardized the CWM values and fitted a students-t distributed likelihood for pollen CWM ($CWM_{mean}^{pol}$) with vegetation CWM ($CWM_{mean}^{veg}$) as the single predictor so that,

$$  CWM_{mean}^{pol} \sim t(\beta_{t} CWM_{mean}^{veg},CWM_{sd}^{pol},df) $$

The slope $\beta_{t}$ is a correlation-like metric defined with a flat, uninformative prior. 
$$
\beta \sim Uniform(-1,1)
$$

We estimated the degrees of freedom ($df$) of the t-distribution with and exponential prior with a mean of 30, the t-distribution becomes almost normal at values of 30 and above (Kruscke, dbda),

$$
df \sim Exp(\lambda)
$$

$$
\lambda = 1/30
$$

### Treatments 
To test the ability of pollen correction factors to improve pollen - vegetation relationships, we divided pollen counts by pollen productivity estimates. We used pollen productivity estimates (ppe's) from @Githumbi_2022 who provides ppe's and uncertainty estimates for 39 taxa in Europe. Pollen productivity estimates were not available for all taxa in the pollen data, taxa without pollen productivity factors were excluded. On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%).

To test for differences in the representation of particular plant functional groups by pollen assemblages, pollen were divided in to plant functional groups. We used the classification of @Reitalu_2019 who distinguished between woody (trees and shrubs) and non-woody species (grasses, herbs and dwarf shrubs of the Ericaceae family). @Reitalu_2019 was also used to classify pollen in to pollination mode, either insect-pollinated or not insect-pollinated. To classify the vegetation in to plant functional groups and pollination mode, data from the TRY database was used [@kattge2020try].

To test for the effect of taxonomic resolution on the trait reconstructions, we calculated CWM trait values of the vegetation using all available trait data of the species, genus or family.

### Community weighted mean changes with elevation
To test if trait composition changes of the vegetation along an elevation gradient are reflected by pollen based trait reconstructions, we fitted a generalized linear model with elevation as the predictor, so that the likelihood of the CWM value is defined as follows,

$$CWM^{mean} \sim t(\alpha + \beta Y_{elev}, CWM^{sd},df)$$

Where $\alpha$ and $\beta$ are the intercept and the slope, respectively, which are supplied by uninformative priors. $Y_{elev}$ is the elevation as the linear predictor. We modeled this relationship for the grassland and forest plots separately.

## Results

### Pollen representation by the vegetation
- missing species in pollen 
- 
### Effect of spatial scale and plant functional group
Over all treatments and spatial scales, we only found evidence of a significant positive relationship between pollen-based reconstructed CWM trait values and vegetation CWM values for plant height. We found little support for a positive association between pollen and vegetation trait value for leaf area and specific leaf area. The spatial scale matters for the relationship between pollen plant height and vegetation plant height, only at scales between 10-100 m the relationship is significant.

Although pollen productivity estimates do seem to improve the representation of the taxa in the vegetation by the pollen data (Supplementary information), it does not help to improve the pollen-vegetation trait relationships. Taking in account uncertainty around pollen productivity estimates, does XXX by between --- and ---

![Estimaed correlation between the pollen data and vegetation data. Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval. Sample size is noted above the mean correlation coefficient.](Figures/Pollen_correction_factors_regression.png)

\* add uncertainty around pollen correction factors

We expected that the trait composition of wind-pollinated and woody vegetation would be better represented by the pollen data, especially at greater spatial scale. The plant height of not wind-pollinated taxa is well represented by the pollen data at the smallest spatial scale (0-10 m), but not at greater spatial scales or for other traits. The subset of wind-pollinated taxa does not show significant relsationships. Sub-setting pollen and vegetation data in to woody and non-woody vegetation did not improve pollen - vegetation trait relationships.

![Estimated correlation between the pollen data and vegetation data. Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval. Sample size is noted above the mean correlation coefficient.](Figures/Pollination_mode_regression.png)

![Estimated correlation between the pollen data and vegetation data. Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval. Sample size is noted above the mean correlation coefficient.](Figures/PFT_regression.png)

We expected pollen to better represent trait composition at higher taxonomic resolution, family or genus level, because of the general low taxonomic resolution of pollen data. We did not find that increasing the taxonomic resolution on which the trait distribution are modelled, improves the relationship between pollen and vegetation trait values.

![Estimated correlation between the pollen data and vegetation data. Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval. Sample size is noted above the mean correlation coefficient.](Figures/Taxonomic_resolution_regression.png)

### CWM change with elevation

We expected to pollen to reproduce similar trends in trait change along an elevation gradient. Pollen do not represent the changes in vegetation CWM values well. Pollen underestimate LA and SLA values and the decrease in LA and SLA in grasslands that was observed in the vegetation is not predicted by the pollen data. Also the steep decrease in plant height along the elevation gradient is not reproduced by the pollen data, and the plant height in grassland is overestimated.

![Vegetation CWM trait change with elevation for leaf area, plant height and specific leaf area and two vegetation types: forest and grassland. Darker shaded area shows the 95% credibility interval and lighter shade area shows the 95% prediction interval.](Figures/CWM_elevation_vegetaion.png)

![Pollen CWM trait change with elevation for leaf area, plant height and specific leaf area and two vegetation types: forest and grassland. Darker shaded area shows the 95% credibility interval and lighter shade area shows the 95% prediction interval.](Figures/CWM_elevation_pollen.png)

## Discussion

Pollen assemblages are the main proxy for studying past vegetation features, and are a reliable proxy for plant species richness and XX. However the relationship with vegetation trait composition is complex. Here we find little evidence for pollen being a reliable proxy for plant trait composition.but the relationship between pollen assemblages and the vegetation composition is complex Exploring the possible causes for the lack of a relationship between pollen CWM values and vegetation CWM values, we suggest that this either has to do with the representation of the vegetation by the pollen data (abundance, species richness) or the large trait variation within pollen types.

vegetation structure (Switzerland) and scale (Scotland)

Previous studies have showed that pollen assemblages are a reliable proxy for plant richness (@Abraham_2022, @Reitalu_2019), but these are reasons why to be cautious. Taxa can become over- or underrepresented in the pollen record because of taxon-specific differences in pollen production. Here we applied pollen productivity estimates to correct for the overproduction of certain taxa. While apply these correction factors improved the representation of the vegetation by the pollen data (supplementary), Applying this correction did not improve the relationship between pollen and vegetation trait values, however. Wind-pollinated taxa and woody taxa are generally best represented by pollen assemblages, but separating the data in to these two categories had no effect.

The spatial scale that is represented by the pollen data is an important factor influencing the representativeness of pollen for vegetation reconstruction. All our treatments were assessed at three spatial scales. It seems that the intermediate scale, up to 100 m, is the most representative for the pollen data. But the relevant spatial scale is also dependent on the openness of the landscape (Bunting, ABraham).

Despite here showing little evidence for pollen being a reliable proxy for past trait composition, we do not disencourage the use of it, we advocate for caution of the use., rather encaurage further testing of the validity of this proxy. And finding ways to improve it. X-Y studies have found a relationship, and looking at the confidence interval of the data we do not disproof the existence of a possitive correlation. We take into account the uncertainty around pollen trait values, but we had a very low sample size and a diverse landscape. Also moss samples vs. lake samples.

the traits of species that do not show up in the dat.a

## Conclusion
Characterizing the factors that influence the relationship between vegetation trait composition and , and validating the relationship between trait composition of the vegetation and is an important step towards the application of traits in palaeoecology.

-   swiss - grass land receiving input from forests

-   traits of species that do not show up in the pollen record.

These studies have showed

-   representativeness of trait databases for trait measurements in the field?

    -   Sandel (2015) - trait databases are biased towards big leaved and big seeded plants
    -   Prinzing (2008) - Less lineages -- more trait variation
    -   Sterwart (2022, preprint) - gapfilling performs well (?)

-   Look for elevation studies in pollen ecology

-   Being careful about apply trait based palaeoecology

    -   bigger sample size

-   Interspecific trait variation in LA

-   positive relationships found by others

-   Some pollen types consider many more species than others for example in this study XX . This is also identified as a problem in plant richness studies (Reitalu 2019) and trait rep (Carvalho, Blaus). although bayesian xxx

-   Relationship is not significant, though still could be positive, as enocmpossed by the 95% confidence interval. More data, less uncertainty in the CWM estimates .

-   size of the relevant source area (Bunting 2003)

    -   The spatial resolution of a pollen assemblage is difficult to assess and is dependent on several factors, such as differences in taxon-specific differences in pollen productivity and dispersal and landscape structure

    -   here we surveyed sites with different openness .

    -   The relationship between spatial scale and pollen records is complex (\@Abraham_2022) and dependant on the species richness and landscape openness

        -   forest data not significantly related to florisitc data while open landscapes are.

        -   highest correlation between pollen and species within 1.5-70 m in open landscapes

        -   high correlation with pollen richness and landscape openess

        -   Changes in the relationship with changing spatial scale are largely driven by the numbers of species newly appearing as the radius of surveyed area increases, especially as new habitats are added (DIRECT COPY)

-   High productivity of tree pollen bias pollen-plant diversity relationships @Odgaard_1999

-   Trees intercept pollen in their canopies @Jackson_1999

-   pollen based trait reconstructions have been shown to correctly reflect environmental gradients as elevation (Sande) and and climatic variations (Barboni)

-   exploring the reasons why we did not find correlations between pollen and vegetation values it could be:

    -   low sample size

    -   low pollen count (not picking up rare pollen in the vegetation) (run CWM without the species in the pollen record)

    -   large trait variation in pollen

    -   Is this a pollen problem or a trait problem?

## References

    <div id="refs"></div>

## Appendix

## Pollen percentages

![Pollen percentage of the most abundant (\> 5%) taxa in Scotland. The effect of the application of pollen productivity estimates to the representation of pollen taxa is shown by the coloured bars. ![Pollen percentage of the most abundant (\> 5%) taxa in Switzerland. Sites are sorted in order of increasing elevation (top to bottom). The effect of the application of pollen productivity estimates to the representation of pollen taxa is shown by the coloured bars.](Figures/Pollen_percentage_swiss.png)](Figures/Pollen_percentage_scotland.png)

## Pollen representation

![Scatter plot of vegetation basal area % against pollen abundance %. Coloured dots represent either wind or non-wind pollinated taxa.](Figures/Pollen_rep_percent.png)

![Scatter plot of vegetation basal area % against pollen abundance % adjusted with pollen productivity estimates. Grey dots show the effect of uncertainty around pollen productivity estimates. Coloured dots represent either wind or non-wind pollinated taxa.](Figures/Pollen_rep_adjustedpercent.png)

### Study design

![Study design in Scotland. The moss sample was collected in the middle of the inner ring. Vegetation surveys were performed in the inner and middle ring. Trait data was collected in the middle ring. The outer ring was surveyed using land cover maps and QGIS](images/paste-941650FE.png){alt="Study design in Scotland. The moss sample was collected in the middle of the inner ring. Vegetation surveys were performed in the inner and middle ring. Trait data was collected in the middle ring. The outer ring was surveyed using land cover maps and QGIS" width="573"}

![Design of Scottish vegetation survey in zone A](Other/Scotland_field_design.png){alt="Design of Scottish vegetation survey in zone A"}

![Design of Swiss trait sampling and vegetation surveys. The moss sample was collected in the centre of the plot.](Other/BDM_study_design.png)

```{r}
bdm_meta <- readRDS("RDS_files/04_zCWM_estimates_vegetation_Switzerland.rds")

ggplot(data = bdm_meta) + 
  geom_point(aes(x = elevation, y = TAP, colour = habitat01)) +
  scale_color_manual("", 
                     values = c(forests = "darkorange", grasslands = "purple"),
                     label = c(forests = "Forests", grasslands = "Grassland")) + 
  scale_x_continuous("Elevation (m.s.l.)") +
  scale_y_continuous("Total annual precipitation (mm)") +
  theme_bw()

ggplot(data = bdm_meta) + 
  geom_point(aes(x = elevation, y = MAT, colour = habitat01)) +
  scale_color_manual("", 
                     values = c(forests = "darkorange", grasslands = "purple"),
                     label = c(forests = "Forests", grasslands = "Grassland")) + 
  scale_x_continuous("Elevation (m.s.l.)") +
  scale_y_continuous("Mean annual temperature (°C)") +
  theme_bw()

```

### Trait data

![Comparison between the trait data collected in the field with trait data obtained from TRY](Figures/LA_trait_values.png)

![Comparison between the trait data collected in the field with trait data obtained from TRY](Figures/PlantHeight_trait_values.png)

![Comparison between the trait data collected in the field with trait data obtained from TRY](Figures/SLA_trait_values.png)

### Taxon level trait values

![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_LA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_SLA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_PH.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_LA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_SLA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_PH.png)

### Scatter plots pollen and vegetation trait values

![](Figures/Scatter_correction.png)

![](Figures/Scatter_growthform.png)

![](Figures/Scatter_pollination.png)

### ![](Figures/Scatter_taxonomic_resolution.png)

### Effect of removal of taxa on CWM trait values

![](Figures/CWM_pollen_removed_taxa.png)